---
title: "Report: `r params$project`"
output:
  rmdformats::robobook:
    self_contained: true
    thumbnails: false
    lightbox: true
    gallery: true
    highlight: tango
    toc_depth: 3
params:
  project: test-gene-level
  date: 2022-05-27
  version: v0.4.2
  regenie_out: ../../../../../results/regenie_step2/pims_step2_Y1.regenie
  regenie_filename: pims_step2_Y1.regenie.all.gz
  phenotype: Y1
  phenotype_file: ../../../../../results/rscript_annotate/pims_phenotype.txt
  pc_plot_file: ../../../../../results/python/draw_pc_plot/pims_pc_plot.png
  eda_plots: ../../../../../results/python/exploratory_data_analysis/*.png
  plot_ylimit: 0
  genes_num: 0
  mask_file: ../../../../../results/rscript_annotate/pims.masks
  manhattan_annotation_enabled: true
  annotation_min_log10p: 1
  r_functions: ./functions.R
  rmd_pheno_stats: ./child_phenostatistics.Rmd
  all_params: ""
---

<!--
Copied from: https://github.com/genepi/nf-gwas/blob/main/reports/gene_level_report_template.Rmd
 Original authors:
    Sebastian Schönherr
    Lukas Forer
    Johanna Schachtl-Riess
    Silvia Di Maio

 Modified by:
    Piotr Suszyński
  
  toc_depth: 3
  fig_width: 16
  fig_height: 10
-->

```{r setup, include=FALSE}
library(DT)
library(data.table)
library(R.utils)
library(ggplot2)
library(skimr)
library(kableExtra)
library(dplyr)
library(tidyr)
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, error = TRUE, cache=FALSE)
source(params$r_functions)
```


# Project summary

| Parameter        | Value                       |
|------------------|-----------------------------|
| Project          | `r params$project`          |
| Pipeline Version | `r params$version`          |
| Date             | `r params$date`             |
| Phenotype File   | `r params$phenotype_file`   |
| PC Plot File     | `r params$pc_plot_file`     |
| EDA Plot Files   | `r params$eda_plots`        |
| Phenotype        | `r params$phenotype`        |
| Mask File        | `r params$mask_file`        |
| Regenie Output   | `r params$regenie_filename` |
| Number of genes  | `r params$genes_num`        |
| Parameters       | `r gsub(", ", "|\n| |", gsub(":", " :  ", gsub("\\|", "\\\\|", params$all_params)))`       |


## Gene-based settings

```{r}
regenieTable <- fread(params$regenie_out, header=TRUE, tmpdir="temp")
regenieTable <- regenieTable %>% tidyr::separate(ID,
                                                 into=c("Set", "Mask", "AAF"),
                                                 sep="\\.",
                                                 extra="merge")
regenieTable$joint <- ifelse(grepl("BURDEN", regenieTable$TEST), "yes", "no")
regenieTable_burdenJoint <- select(filter(regenieTable, joint == "yes"), c(CHROM:Set,TEST,CHISQ:joint))
regenieTable <- filter(regenieTable, joint == "no")
regenieTable <- regenieTable %>% tidyr::unite(.,
                                              "Settings",
                                              TEST, AAF, Mask,
                                              sep = ".",
                                              remove = FALSE)
```

### Tests

```{r}
gene_tests <- unique(regenieTable$TEST)
kable(gene_tests,
      col.names = c("Tests"))
```


### AAF cutoffs

```{r}
aaf_cutoffs <- unique(regenieTable$AAF)
kable(aaf_cutoffs,
      col.names = c("AAF cutoffs"))
```

### Masks
```{r}
masks <- read.table(params$mask_file, header=FALSE)
names(masks) <- c("Mask", "Annotations")
masks$Annotations <- gsub(",", ", ", masks$Annotations)
kable(masks)
```


# Input data characteristics

## Phenotype Statistics

```{r child=params$rmd_pheno_stats}

```

## Exploratory Data Analysis

(click to expand)

```{r, echo=FALSE, out.width="50%", out.height="50%", fig.cap="caption", fig.show='hold', fig.align='center'}
file_list <- unlist(strsplit(params$eda_plots, "\\s+"))
file_list <- sort(file_list[file_list != ""])

knitr::include_graphics(file_list)
```

## PC Plot

```{r, echo=FALSE, fig.align='center'}
knitr::include_graphics(params$pc_plot_file)
```

# Association tests results

### Joint tests for burden masks
https://github.com/genepi/nf-gwas/blob/main/reports/functions.R
```{r, results='asis'}
if(nrow(regenieTable_burdenJoint) > 0) {
    test_joint <- unique(regenieTable_burdenJoint$TEST)
    test_joint <- as.data.frame(test_joint)
    test_joint <- test_joint %>% tidyr::separate(test_joint, into=c("Model", "Test", "Joint"), sep="-", extra="merge")
    cat("The following joint tests were used for burden masks:", unique(test_joint$Joint), ".")
} else {
    cat("No joint test was performed for burden masks and AAF thresholds.")
}

```


```{r, eval=nrow(regenieTable_burdenJoint) > 0, results='asis', dpi=300}
cat('\n\n')
cat('## Joint test(s) for burden masks and AAF thresholds')
cat('\n\n')
if(params$plot_ylimit == 0) {
  limit = NULL
  } else {
    limit = c(0, params$plot_ylimit )
    cat(paste("Please note that the Y-axis limit has been set to '", params$plot_ylimit,"'. Please use the 'plot_ylimit' parameter to adapt this behaviour."))
  }

for (test in unique(regenieTable_burdenJoint$TEST)){
  regenieTable_tmp01 <- regenieTable_burdenJoint[regenieTable_burdenJoint$TEST == test, ]
  cat('\n\n')
  cat('### Test: ', unique(regenieTable_tmp01$TEST))
  cat('\n\n')
  #manhattan plot
  cat('\n\n')
  cat('**Manhattan Plot**')
  cat('\n\n')
  df_prep <- manhattan_prep(regenieTable_tmp01)
  plot <- manhattan_plot(df_prep, ymax = params$plot_ylimit, hlines = FALSE, n_genes = params$genes_num)
  #annotatate plot (if it is enabled)
  if(!params$manhattan_annotation_enabled) {
    plot
    } else {
      plot <- plot + ggrepel::geom_text_repel(data = df_prep %>% group_by(Set) %>%
                             top_n(1, LOG10P) %>%
                               slice(1) %>%
                             filter(LOG10P > params$annotation_min_log10p),
                            aes(label=Set), color = 1, fontface = "italic")
    }
  print(plot)
  #qqplot
  cat('\n\n')
  cat('**QQ Plot**')
  cat('\n\n')
  pp <- qq_plot(df=regenieTable_tmp01)
  print(pp)
}

```


```{r, results='asis', dpi=300}
for (test in unique(regenieTable$TEST)){
  regenieTable_tmp01 <- regenieTable[regenieTable$TEST == test, ]
  cat('\n\n')
  cat('## Test: ', unique(regenieTable_tmp01$TEST))
  cat('\n\n')
  for (settings in unique(regenieTable_tmp01$Settings)) {
    regenieTable_tmp02 <- regenieTable_tmp01[regenieTable_tmp01$Settings == settings, ]
    annotations <- masks[masks$Mask == unique(regenieTable_tmp02$Mask),]$Annotations
    annotations <- gsub(",", ", ", annotations)
    cat('\n\n')
    cat('### AAF cutoff: ', unique(regenieTable_tmp02$AAF), ', Mask: ', unique(regenieTable_tmp02$Mask), ' (', annotations, '), (Test: ', unique(regenieTable_tmp02$TEST), ')',
        sep = "")
    cat('\n\n')
    #note for manual ylim setting (if ymax is defined)
    if(params$plot_ylimit == 0) {
      limit = NULL
      } else {
        limit = c(0, params$plot_ylimit )
        cat(paste("Please note that the Y-axis limit has been set to '", params$plot_ylimit,"'. Please use the 'plot_ylimit' parameter to adapt this behaviour."))
      }
    #manhattan plot
    cat('\n\n')
    cat('**Manhattan Plot**')
    cat('\n\n')
    df_prep <- manhattan_prep(regenieTable_tmp02)
    plot <- manhattan_plot(df_prep, ymax = params$plot_ylimit, hlines = FALSE, n_genes = params$genes_num)
    #annotatate plot (if it is enabled)
    if(!params$manhattan_annotation_enabled) {
      plot
      } else {
        plot <- plot + ggrepel::geom_text_repel(data = df_prep %>% group_by(Set) %>%
                             top_n(1, LOG10P) %>%
                               slice(1) %>%
                             filter(LOG10P > params$annotation_min_log10p),
                            aes(label=Set), color = 1, fontface = "italic")
      }
    print(plot)
    #qqplot
    cat('\n\n')
    cat('**QQ Plot**')
    cat('\n\n')
    pp <- qq_plot(df=regenieTable_tmp02)
    print(pp)
  }
}
```


------------------------------------------------------------------------

<small>
This report has been created with **[rare-var-assoc-nf `r params$version`](https://github.com/biodatageeks/rare-var-assoc-nf)**, a nextflow pipeline developed by [Piotr Suszyński](mailto:piotr.suszynski.dokt@pw.edu.pl) and [Tomasz Gambin](mailto:tomasz.gambin@pw.edu.pl), Politechnika Warszawska. The Manhattan and QQ plots reporting was adapted from **[nf-gwas `r params$version`](https://github.com/genepi/nf-gwas)**, a nextflow pipeline developed by [Sebastian Schönherr](mailto:sebastian.schoenherr@i-med.ac.at), [Lukas Forer](mailto:lukas.forer@i-med.ac.at) and [Johanna Schachtl-Riess](mailto:johanna.schachtl-riess@i-med.ac.at), Institute of Genetic Epidemiology, Medical University of Innsbruck.
</small>
